 

### 2021-07-27 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

**Leaf photosynthetic thermal tolerance**
Irreversible damage to photosystem II, leading to leaf necrosis and eventually leaf death, occurs at $T_{leaf}$ of $40-60^{\circ}C$ [Baker 2008; @feeley_thermal_2020]
$T_{crit}$ is the temperature at which efficiency of photosystem II starts to decrease, based on chlorophyll a fluorescence; and $T_{50}$ represents temperature at which the efficiency of photosystem II had decreased by 50% [@slotLeafHeatTolerance2020].

**Thermal tolerance (e.g., $T_{50}$) varies with exposure, but there are no studies isolating effect of height.**

*If we take geographic gradients as a reliable proxy, we'd expect temperature sensitivities to reflect the environment to which the trees are adapted/ acclimatized.* Global $T_{50s}$ across latitude and climate range from ~45-50$^\circ$C [@osullivanThermalLimitsLeaf2017], where $T_{50}$ and $T_{crit}$ decrease with elevation among tropical forests and across latitudes [@feeley_thermal_2020;  @slotLeafHeatTolerance2020; @osullivanThermalLimitsLeaf2017].

>>>>

**Leaf photosynthetic thermal tolerance**
Irreversible damage to photosystem II, leading to leaf necrosis and eventually leaf death, occurs at $T_{leaf}$ of $40-60^{\circ}C$ [@bakerChlorophyllFluorescenceProbe2008; @feeley_thermal_2020].Thermal tolerance is described in terms of the temperature at which efficiency of photosystem II starts to decrease, $T_{crit}$, and the $T_{leaf}$ at which efficiency of photosystem II had decreased by 50%, $T_{50}$ [@slotLeafHeatTolerance2020]. Thermal tolerance varies globally across latitude and climate, where $T_{50}$ and $T_{crit}$ decrease across latitudes, and with elevation among tropical forests [@feeley_thermal_2020;  @slotLeafHeatTolerance2021; @osullivanThermalLimitsLeaf2017].
